---
title: "Getting Started with binomial"
author: "Masha Paley"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to binomial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = F, message = F}
knitr::opts_chunk$set(collapse = T, comment = "#>")
library(binomial)
```

### Binomial Distribution

The package `"binomial"` is an implementation of functions calculating probabilities of a binomial random variable as well as related calculations such as the binomial probability distribution, expectation, variance, etc. The package also implements functions for the visualization of the binomial probability distribution.

### Binomial Coefficient

The first function of the `"binomial"` package calculates the binomial coefficient. This coefficient is equal to the number of combinations in which k successes can occur in n trials, sometimes referred to as "n choose k". This binomial coefficient can be calculated using the function `bin_choose()`:

```{r}
bin_choose(5, 2)
```

The output above is equal to the number of combinations in which 2 successes can occur in 5 trials

### Binomial Probability

The probability of k successes in n trials with a given probability of success can be calculated with the function `bin_probability()`. The function takes three arguments:

- `success`: a sinlge number or vector of successful trials 
- `trials`: the number of trials
- `prob`: the probability of a successful trial

```{r}
bin_probability(success = 2, trials = 5, prob = 0.5)
```

The output above is equal to the probability of getting 2 successes in 5 trials, assuming the probability of a successful trial is 0.5

### Binomial Distribution

The binomial probability distribution of a binomial random variable with a set number of trials and probability of a success can be displayed using the function `bin_distribution()`:

```{r}
bin_distribution(trials = 5, prob = 0.5)
```

By default, the distribution is displayed as a data frame with one column for the number of successes and another with the corresponding binomial probability.

### Plotting the Binomial Distribution

The graph of a barplot displaying the probability histogram of a binomial distribution can be displayed using the function `plot()`: 

```{r}
dis1 <- bin_distribution(trials = 5, prob = 0.5)
plot(dis1)
```

### Cumulative Binomial Distribution

The cumulative binomial probability distribution of a binomial random variable with a set number of trials and probability of success can be displayed using the function `bin_cumulative()`:

```{r}
bin_cumulative(trials = 5, prob = 0.5)
```

### Plotting the Cumulative Distribution

Similarly to the simple binomial probability distribution, the cumulative distribution can be plotted using the same `plot()` function:

```{r}
dis2 <- dis2 <- bin_cumulative(trials = 5, prob = 0.5)
plot(dis2)
```

### Creating a Binomial Random Variable

In order to access the summary functionalities of the `"binomial"` package, a `"binvar"` object must first be created using the `bin_variable()` function. The function takes two arguments:

- `trials`: the number of trials
- `prob`: the probability of a successful trial

```{r}
bin_variable(trials = 10, p = 0.3)
```

The above output is of class `"binvar"` and displays a nice table of the binomial random variable's parameters

### Binomial Random Variable Summary

A nice table of variable parameters and measures can be displayed using the `summary()` function which takes a `"binvar"` object as its' argument:

```{r}
var1 <- bin_variable(trials = 10, p = 0.3)
summary(var1)
```

### Calculating Individual Measures

Each of the measures in displayed in the `summary()` output can be calculated invidiually using the following measure functions:

- `bin_mean()`
- `bin_variance()`
- `bin_mode()`
- `bin_skewness()`
- `bin_kurtosis()`

Each function takes two arguments:

- `trials`: the number of trials
- `prob`: the probability of a successful trial

For example, the variance of a binomial random variable with 10 trials and probability of success equal to 0.3 can be calculated like such:

```{r}
bin_variance(10, 0.3)
```

